Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-05-14 ms.tex
more clean-up of low-order correction discussion

The need for additional low-order corrections beyond simple scalar offsets 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a low-level dipole pattern, with an orientation 
consistent with that of the Solar dipole. Our strategy will be to 
simultaneously fit the residual dipole and zero-level offset amplitudes. To 
most precisely recover these amplitudes, it is necessary to have the highest 
available S/N in the independent variable of our regression. For this reason we
have used 857 GHz as a reference for the 100-545 GHz bands, as opposed to the 
FDS99 predictions or H\,\textsc{i} data.

We have performed one regression per HFI band to simultaneously fit for the 
zero level offset, the slope relative to 857 GHz, and the residual dipole 
amplitude. For each 100-545 GHz HFI band, we restrict to regions of low column 
density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ 
km\,s$^{-1}$) and fit the following model:

@@

With $p$ denoting a single nside=64 HEALPix pixel \citep{healpix} in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission and compact sources removed. 
In the context of Equation \ref{equ:dip}, $\nu_i \in$  
\{100, 143, 217, 353, 545\} GHz. Note that $\mathcal{I}_{\nu_i}$ is always in
the native units of the relevant \PLANCK~band. $\mathcal{D}$ is a scaling of 
the Solar dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit amplitude. Because 
$\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we have an 
overconstrained linear model with three parameters: $m$, $d$, and $b$. $m$ 
represents the best-fit slope of \PLANCK~band $\nu_i$ versus \PLANCK~857 GHz 
assuming they are linearly related. $d$ gives the residual Solar dipole
amplitude, and its best-fit value represents the scaling of the Solar dipole 
that makes the \PLANCK~band $\nu_i$ versus 857 GHz correlation most tightly 
linear. $b$ represents the constant offset that must be subtracted from the 
band $\nu_i$ map to make its zero level consistent with that of the 857 GHz 
map. For each band $\nu_i$, we obtain estimates of $m$, $d$, and $b$ by 
performing a linear least squares fit with uniform weights and iterative 
outlier rejection. Figure \ref{fig:dip} shows scatterplots of the band $\nu_i$ 
versus 857 GHz correlation before and after correcting for the best-fit offset 
and residual dipole, for each $\nu_i \in$ \{143, 217, 353, 545\} GHz. Not only 
are the tightened correlations striking in these scatterplots, but the residual
dipole subtractions appear very successful in the two-dimensional band $\nu_i$ 
maps themselves. Before performing thermal dust fits, we therefore subtract the
best-fit $b$ and $d\mathcal{D}$ from each band $\nu_i$ map. The best-fit
offsets and residual dipole amplitudes are listed in Table \ref{table:offs},
along with other important per-band parameters, such as the fractional 
multiplicative calibration uncertainty $c_{\nu}$.

@@

Where $\theta_p$ is the angle of pixel $p$ relative to the Solar velocity 
vector. It is not clear physically why such a term should improve the 
correlation between $\mathcal{I}_{100}$ and $\mathcal{I}_{857}$. However, the 
scatter plots shown in Figure \ref{fig:quad} make it clear that such a term, 
with best-fit amplitude $q$=$-7.17$$\times$10$^{-6}$, dramatically improves the
quality of the linear correlation between the 100 GHz and 857 GHz channels. We 
tried re-running the fits at 143-545 GHz including the $Q_p$ quadrupole term, 
but in all cases found the best-fit quadrupole to be negligible, and yield no 
substantial tightening of the correlation versus 857 GHz.

>>>>

The need for additional low-order corrections beyond simple scalar offsets 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a low-level dipole pattern, with an orientation 
consistent with that of the Solar dipole. Our strategy will be to 
simultaneously fit both the residual dipole and zero-level offset amplitudes. 
To most precisely recover these amplitudes, it is necessary to have the highest
available S/N in the independent variable of our regression. For this reason we
have used 857 GHz as a reference for the 100-545 GHz bands, as opposed to the 
FDS99 predictions or H\,\textsc{i} data.

Thus, we perform one regression per HFI band to simultaneously fit for the 
zero level offset, the slope relative to 857 GHz, and the residual dipole 
amplitude. For each 100-545 GHz HFI band, we restrict to regions of low column 
density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ 
km\,s$^{-1}$) and fit the following model:

@@

With $p$ denoting a single nside=64 HEALPix pixel \citep{healpix} in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission and compact sources removed. 
In the context of Equation \ref{equ:dip}, $\nu_i \in$  
\{100, 143, 217, 353, 545\} GHz. Note that $\mathcal{I}_{\nu_i}$ is always in
the native units of the relevant \PLANCK~band. $\mathcal{D}$ is a scaling of 
the Solar dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit amplitude. Because 
$\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we have an 
overconstrained linear model with three parameters: $m$, $d$, and $b$. $m$ 
represents the best-fit slope of \PLANCK~band $\nu_i$ versus \PLANCK~857 GHz 
assuming they are linearly related. $d$ is the residual Solar dipole amplitude,
and its best-fit value represents the scaling of the Solar dipole that makes 
the \PLANCK~band $\nu_i$ versus 857 GHz correlation most tightly linear. $b$ 
represents the constant offset that must be subtracted from the band $\nu_i$ 
map to make its zero level consistent with that of the 857 GHz map. For each 
band $\nu_i$, we obtain estimates of $m$, $d$, and $b$ by performing a linear 
least squares fit with uniform weights and iterative outlier rejection. Figure 
\ref{fig:dip} shows scatterplots of the band $\nu_i$ versus 857 GHz correlation
 before and after correcting for the best-fit offset and residual dipole, for 
each $\nu_i \in$ \{143, 217, 353, 545\} GHz. Not only are the tightened 
correlations striking in these scatterplots, but the residual dipole 
subtractions appear very successful in the two-dimensional band $\nu_i$ maps 
themselves. Before performing thermal dust fits, we therefore subtract the 
best-fit $b$ and $d\mathcal{D}$ from each band $\nu_i$ map. The best-fit 
offsets and residual dipole amplitudes are listed in Table \ref{table:offs},
along with other important per-band parameters, such as the fractional 
multiplicative calibration uncertainty $c_{\nu}$.

@@

Where $\theta_p$ is the angle of pixel $p$ relative to the Solar velocity 
vector. It is not clear physically why such a term should improve the 
correlation between $\mathcal{I}_{100}$ and $\mathcal{I}_{857}$. However, the 
scatter plots shown in Figure \ref{fig:quad} make it clear that such a term 
dramatically improves the quality of the linear correlation between the 100 GHz
and 857 GHz channels. We tried re-running the fits at 143-545 GHz including the
$Q_p$ quadrupole term, but in all cases found the best-fit quadrupole to be 
negligible, and yield no substantial tightening of the correlation versus 857 
GHz.
